---
title: "Assignment 4 - Coordinating Heart Rate"
author: "Riccardo Fusaroli"
date: "November 6, 2017"
output: html_document
---

Notes from class:
Preprocess
Artifacts = mean of outliers
Downsample = 1 data point for every 100
Scale = changes the scale so they are identical and divide by sd. You can compare how much they vary from their own signal, how many sd
Write 2-3 lines about how you did it

Coordination
There will be some similaities between particpants due to physiology
Shuffled baseline; is it different from the baseline = coordinated due to something else.
Surrogate pair: matching individuals from different trials with each other

Conditions
We manipulated
Inspiring : inhaling
Have the conditions made a difference
Hr as a function of condition

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysing Heart Rate and Respiration data

The goal of this assignment is to first familiarize you with heart rate, and respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.

2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

3) Do you observe differences in coordination between conditions? Report the models and results.

4) Is respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.

N.B. to give you a bit more data I included data from last year (Study1) and from your class (Study2). Note that synchronouns and turn-taking are the same across both studies, but the third condition is different: last year it was self-paced joint reading; this year it was the tv-series conversation. So you might want to exclude the self-paced reading (but, up to you!)

## Step by step suggestions to solve the assignment

### Exploring physiological signals

- Choose one pair (one pair, three conditions)
- Load the logs
- Produce a plot of the participants' respiration signal and a different one of the participants' HR signal (for inspecting whether the data is usable)
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3)
- Can you eye-ball which condition if any displays more physiological coordination?

- Run crqa on heart rate and respiration data (find parameters, run crqa)
- Does this tell you more than just eyeballing the plots?

```{r}
#Set working directory and load data
setwd("~/Desktop/Experimental methods 3/Assignment 4")

data <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/Study2_G5_T1_Synchronous.csv")

library(dplyr)
library(ggplot2)
library(groupdata2)

#Plot before preprocessing
plot1 = ggplot(data)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
plot1

plot2 = ggplot(data)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
plot2

#Preprocessing
#Artifacts = mean of outliers
removeOuts <- function(ts,threshold){
  ts[ts > (mean(ts, na.rm = T) + (threshold * sd(ts, na.rm = T))) |
       ts < (mean(ts, na.rm = T) - (threshold * sd(ts, na.rm = T)))] = mean(ts, na.rm =
       T)  
return(ts)}
threshold=2.5

#remove outliers in condition HR1
plot(data$HR1)
data$HR1=removeOuts(data$HR1,threshold)
plot(data$HR1)

#remove outliers in condition HR2
plot(data$HR2)
data$HR2=removeOuts(data$HR2,threshold)
plot(data$HR2)

#remove outliers in condition Resp1
plot(data$Resp1)
data$Resp1=removeOuts(data$Resp1,threshold)
plot(data$Resp1)

#remove outliers in condition Resp2
plot(data$Resp2)
data$Resp2=removeOuts(data$Resp2,threshold)
plot(data$Resp2)

# Downsample = 1 data point for every 100
d1 = data %>%
    group(n = 100, method = 'greedy') %>%
    dplyr::summarise(
       time = mean(time,na.rm=T),
       HR1 = mean(HR1,na.rm=T),
       HR2 = mean(HR2,na.rm=T),
       Resp1 = mean(Resp1,na.rm=T),
       Resp2 = mean(Resp2,na.rm=T))


# Scale = changes the scale so they are identical and divide by sd. You can compare how much they vary from their own signal, how many sd
d1$Resp1S=scale(d1$Resp1)  
d1$Resp2S=scale(d1$Resp2)  
d1$HR1S=scale(d1$HR1)
d1$HR2S=scale(d1$HR2)

#plots after preprocessing
plot3 = ggplot(d1)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
plot3

plot4 = ggplot(d1)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
plot4

#Overview of all plots
gridExtra::grid.arrange(plot1, plot2, plot3, plot4, ncol=2, nrow=2)

#Run crqa on heart rate and respiration data (find parameters, run crqa)
#Does this tell you more than just eyeballing the plots?
library(crqa)

#crqa resp
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 

ans_resp = optimizeParam(d1$Resp1S, d1$Resp2S, par, min.rec= 3.5, max.rec= 4.5)

#Choose a common value for delay, emb.dim and radius 
Results_resp=crqa (d1$Resp1S, d1$Resp2S, delay=ans_resp$delay, embed=ans_resp$emddim, radius=ans_resp$radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2) 

Results_resp

#crqa heart
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 150,  radiussample = 20, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 

ans_hr = optimizeParam(d1$HR1S, d1$HR2S, par, min.rec= 3.5, max.rec= 4.5)

#Choose a common value for delay, emb.dim and radius 
Results_hr=crqa (d1$HR1S, d1$HR2S, delay=ans_hr$delay, embed=ans_hr$emddim, radius=ans_hr$radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2) 

Results_hr

#Reccurence plot
RP=Results_resp$RP 
RP = matrix(as.numeric(RP), nrow = ncol(RP))  
cols = c("white","blue4") 
image(RP, xlab = "", ylab = "", col = cols) 


RP=Results_hr$RP 
RP = matrix(as.numeric(RP), nrow = ncol(RP))  
cols = c("white","blue4") 
image(RP, xlab = "", ylab = "", col = cols) 


#Error ok; code is working
#Optimal Radius Not found: try again choosing a wider radius span and larger sample sizeFejl i if (length(ts1) < embed * delay) { : argument har længden nul

#To check if the code works, insert values in delay, embed and radius

```
[Write 2-3 lines about how you did it.
I did preprocessing which consisted of downsampling, removal of outliers, rescaling the signal, eyeballing the signal and doing a recurrence plot. 

Can you eye-ball which condition if any displays more physiological coordination?
When you just eyeball the plots it is difficult to see the which condition displays more physiological coordination due to the large amount of signal. The recurrence plot displays only recurrence and highlight the physiological coordination, which makes it easier to spot.]

### Systematically pre-process the data
- Loop through all the files (either with a loop or with a function), check which files should be excluded, if any, and save the pre-processed time-series. Tip: plot and visually inspect the data to figure out which should be excluded.
- Run crqa on all the pre-processed time-series and save the output (don't forget to add columns with study, group, condition and trial). Tip: remember to first assess optimal parameters (dimensions, delay, radius) across all timeseries. Tip: it will often fail, just take whatever parameters you get, select optimal across timeseries parameters and run crqa on all timeseries with those. Tip: double check the rr. When I ran the loop, I got very low rr, so I adjusted the radius until the average of rr across all pairs was approx. 4%.

```{r}

library(readr)

#loop 1 - Preprocessing
files = list.files(pattern = "csv")


for (i in files){
  print(i)
  data = read.csv(i) 
 
  #downsampling
  d1 = data %>%
    group(n = 100, method = 'greedy') %>%
    dplyr::summarise(
       time = mean(time,na.rm=T),
       HR1 = mean(HR1,na.rm=T),
       HR2 = mean(HR2,na.rm=T),
       Resp1 = mean(Resp1,na.rm=T),
       Resp2 = mean(Resp2,na.rm=T))
 
  # remove outliers
  d1$HR1=removeOuts(d1$HR1,threshold)
  d1$HR2=removeOuts(d1$HR2,threshold)
  d1$Resp1=removeOuts(d1$Resp1,threshold)
  d1$Resp2=removeOuts(d1$Resp2,threshold)
  
  #rescale
  d1$Resp1S=scale(d1$Resp1)  
  d1$Resp2S=scale(d1$Resp2)  
  d1$HR1S=scale(d1$HR1)
  d1$HR2S=scale(d1$HR2)
  
  
  #save in new folder
  file1 = paste0("pp_", i)
  write_csv(d1, file.path("pp",file1))
  
}
```

Plot all data, which should be excluded?

```{r}
#ret working directory
dp1 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G1_T1_Synchronous.csv")
dp2 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G1_T2_TurnTaking.csv")
dp3 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G2_T1_TurnTaking.csv")
dp4 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G2_T2_Synchronous.csv")
dp5 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G3_T2_TurnTaking.csv")
dp6 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G3_T3_Synchronous.csv")
dp7 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G4_T1_Synchronous.csv")
dp8 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G4_T3_TurnTaking.csv")
dp9 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study2_G10_T1_Conversation.csv")

#Not good - get rid of data 
pResp1 = ggplot(dp1)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp1

#Not good - get rid of data 
pResp2 = ggplot(dp2)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp2

#Not good - get rid of data 
pResp3 = ggplot(dp3)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp3

#Not good - get rid of 
pResp4 = ggplot(dp4)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp4

dpHR10 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G10_T2_TurnTaking.csv")
pHR10 = ggplot(dpHR10)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR10

dpResp10 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G10_T2_TurnTaking.csv")
pResp10 = ggplot(dpResp10)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp10


dpHR11 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G10_T3_Synchronous.csv")
pHR11 = ggplot(dpHR11)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR11

dpResp11 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G10_T3_Synchronous.csv")
pResp11 = ggplot(dpResp11)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp11

dpHR12 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T1_Synchronous.csv")
pHR12 = ggplot(dpHR12)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR12

dpResp12 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T1_Synchronous.csv")
pResp12 = ggplot(dpResp12)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp12

dpHR13 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T2_TurnTaking.csv")
pHR13 = ggplot(dpHR13)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR13

dpResp13 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T2_TurnTaking.csv")
pResp13 = ggplot(dpResp13)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp13


dpHR14 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T3_Conversation.csv")
pHR14 = ggplot(dpHR14)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR14

dpResp14 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T3_Conversation.csv")
pResp14 = ggplot(dpResp14)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp14

dpHR15 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T1_TurnTaking.csv")
pHR15 = ggplot(dpHR15)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR15

dpResp15 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T1_TurnTaking.csv")
pResp15 = ggplot(dpResp15)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp15

dpHR16 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T2_Conversation.csv")
pHR16 = ggplot(dpHR16)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR16

dpResp16 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T2_Conversation.csv")
pResp16 = ggplot(dpResp16)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp16

dpHR17 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T3_Synchronous.csv")
pHR17 = ggplot(dpHR17)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR17

dpResp17 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T3_Synchronous.csv")
pResp17 = ggplot(dpResp17)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp17

dpHR18 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G7_T1_Conversation.csv")
pHR18 = ggplot(dpHR18)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR18

dpResp18 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G7_T1_Conversation.csv")
pResp18 = ggplot(dpResp18)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp18

dp19 = read_csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G7_T2_Synchronous")

pResp19 =ggplot(dp19)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp19

pHR19 =ggplot(dp19)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR19

dp20 = read_csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G7_T3_TurnTaking")

pResp20 =ggplot(dp20)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp20

pHR20 =ggplot(dp20)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR20

dp21 = read_csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G8_T1_TurnTaking")

pResp21 =ggplot(dp21)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp21

pHR21 =ggplot(dp21)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR21

dp22 = read_csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G8_T2_Synchronous")

pResp22 =ggplot(dp22)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp22

pHR22 =ggplot(dp22)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR22

dp23 = read_csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G8_T3_Conversation")

pResp23 =ggplot(dp23)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp23

pHR23 =ggplot(dp23)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR23

dp24 = read_csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G9_T1_Synchronous")

pResp24 =ggplot(dp24)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp24

pHR24 =ggplot(dp24)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR24

dp25 = read_csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G9_T2_Conversation")

pResp25 =ggplot(dp25)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp25

pHR25 =ggplot(dp25)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR25

dp26 = read_csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G9_T3_TurnTaking")

pResp26 =ggplot(dp26)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp26

pHR26 =ggplot(dp26)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR26

#5 files have been excluded

```

Loop through all files(minus the excluded) to calculate the optimazed parameters for both resp and hr

```{r}

#function fitCRQA

fitCrqa <- function(ts1,ts2){

  par = list(lgM =  30, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 100, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")

  Pars = try(optimizeParam(ts1, ts2, par, min.rec = 2, max.rec = 8))

  if (length(Pars)>1){
    Results=crqa(ts1, ts2, delay=Pars$delay, embed=Pars$emddim, radius=Pars$radius, normalize=0, rescale=0, mindiagline = 2, minvertline = 2)
  } else {
    Results=NULL
    Pars=NULL
    Pars$delay=NA
    Pars$emddim=NA
    Pars$radius=NA
    Results$RR=NA
    Results$DET=NA
    Results$maxL=NA
    Results$L=NA
    Results$ENTR=NA
    Results$LAM=NA
    Results$TT=NA
  }

  df=data.frame(Delay=Pars$delay,Embed=Pars$emddim,Radius=Pars$radius,RR=Results$RR,DET=Results$DET,LMAX=Results$maxL,L=Results$L,ENTR=Results$ENTR,LAM=Results$LAM,TT=Results$TT)

  return(df)

}

library(stringr)

# KØR DET HELE SAMMEN 
setwd("~/Desktop/Experimental methods 3/Assignment 4/pp")

files = list.files(pattern = "pp_")

for (i in files){
  print(i)
  d = read.csv(i)
  resp_df = fitCrqa(d$Resp1S,d$Resp2S)
  heart_df = fitCrqa(d$HR1S,d$HR2S)
  
  Study = str_extract(i,"Study+\\d+")
  Group = str_extract(i,"G+\\d+")
  Trial = str_extract(i,"T+\\d+")
  Condition = substr(i, start = 17, stop = 20)
  
  df = data.frame(Study, Group, Trial, Condition,
                  RespDelay=resp_df$Delay, RespEmbed = resp_df$Embed, RespRadius = resp_df$Radius, 
                  RespRR=resp_df$RR, RespDET = resp_df$DET, RespMaxL = resp_df$LMAX,RespL=resp_df$L, 
                  RespEntr = resp_df$ENTR, RespLAM = resp_df$LAM, RespTT=resp_df$TT, 
                  HeartDelay = heart_df$Delay, HeartEmbed = heart_df$Embed, HeartRadius = heart_df$Radius,
                  HeartRR = heart_df$RR, HeartDET = heart_df$DET, HeartMaxL = heart_df$LMAX, 
                  HeartL = heart_df$L, HeartEntr = heart_df$ENTR, HeartLAM = heart_df$LAM, 
                  HeartTT = heart_df$TT)
                  
  if(exists("dataframe")){dataframe = rbind(dataframe, df)} else {dataframe = df}
  
}

```

CRQA loop through all files with the median of the optimized paramters
```{r}

#median of optimized parameters
hr_delay= round(median(dataframe$HeartDelay, na.rm=T),2) 
hr_embed= round(median(dataframe$HeartEmbed, na.rm=T),2)
hr_radius=round(median(dataframe$HeartRadius, na.rm=T),2) + 1.2 
# 1.2 added to hr_radius; Riccardo figured this out through trial and error

resp_delay= round(median(dataframe$RespDelay, na.rm=T),2)
resp_embed= round(median(dataframe$RespEmbed, na.rm=T),2)
resp_radius=round(median(dataframe$RespRadius, na.rm=T),2)

#crqa loop

setwd("~/Desktop/Experimental methods 3/Assignment 4/pp")

files1 = list.files(path = ("~/Desktop/Experimental methods 3/Assignment 4/pp"), pattern = "pp")

#Making lists for the loop to fill with the relevant parameters for heart rate and respiration 
HR_RR = NULL
HR_DET = NULL
HR_maxL = NULL
HR_L = NULL
HR_ENTR = NULL
HR_LAM = NULL
HR_TT = NULL

Resp_RR = NULL
Resp_DET = NULL
Resp_maxL = NULL
Resp_L = NULL
Resp_ENTR = NULL
Resp_LAM = NULL
Resp_TT = NULL
n = 1

#Loop to run crqa on all time-series and saving the output in the premade lists 
for (i in files1){
  
  file = read.csv(i)
  
  print(i)
  
  #Extracting study, pair, trial and condition from filenames
  Study[n] = str_extract(i,"Study+\\d+")
  Group[n] = str_extract(i,"G+\\d+")
  Trial[n] = str_extract(i,"T+\\d+")
  Condition[n] = substr(i, start = 17, stop = 20)
  
#Heart rate
  results = try(crqa(file$HR1S, file$HR2S, delay = hr_delay, embed = hr_embed, radius= hr_radius, normalize=0,rescale=0, mindiagline = 2, minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(results) > 1) {
    HR_RR[n] <- results$RR
    HR_DET[n] <- results$DET
    HR_maxL[n] <- results$maxL
    HR_L[n] <- results$L
    HR_ENTR[n] <- results$ENTR
    HR_LAM[n] <- results$LAM
    HR_TT[n] <- results$TT
  }
  
#Respiration
    results = try(crqa(file$Resp1S, file$Resp2S, delay = resp_delay, embed = resp_embed, radius= resp_radius, normalize=0,rescale=0, mindiagline = 2, minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(results) > 1) {
    Resp_RR[n] <- results$RR
    Resp_DET[n] <- results$DET
    Resp_maxL[n] <- results$maxL
    Resp_L[n] <- results$L
    Resp_ENTR[n] <- results$ENTR
    Resp_LAM[n] <- results$LAM
    Resp_TT[n] <- results$TT
  }
  n= n+1
}

#Making a dataframe with the relevant parameters from the crqa 
df1 = data.frame(Study, Group, Trial, Condition, HR_RR, HR_DET, HR_maxL, HR_L, HR_ENTR, HR_LAM, HR_TT, Resp_RR, Resp_DET, Resp_maxL, Resp_L, Resp_ENTR, Resp_LAM, Resp_TT)

```
### Creating controls: shuffled controls
 - loop through all pairs and conditions
 - shuffle the timeseries (take a timeseries and rearrange its values in a random order). Tip check the sample() function
 - run crqa and save the output. NB. which delay, embed, radius parameters should you use?
 - statistically compare the crqa indexes in real and shuffled pairs
 
```{r}

setwd("~/Desktop/Experimental methods 3/Assignment 4/pp")

files1 = list.files(path = ("~/Desktop/Experimental methods 3/Assignment 4/pp"), pattern = "pp")

for (i in files1){
  
  file = read.csv(i)
  
  print(i)
  
  shuffleddata = file[sample(1:nrow(file), 2401, replace = T ), ]
  
  file2 = paste0("shu_", i)
  write_csv(shuffleddata, file.path("shuffleddata", file2))
  
}


```


Loop to run crqa on the shuffled timeseries and saving the output 
```{r}

setwd("~/Desktop/Experimental methods 3/Assignment 4/pp/shuffleddata")
files2 = list.files(path = ("~/Desktop/Experimental methods 3/Assignment 4/pp/shuffleddata"), pattern = "shu")
files2

#Making lists for the crqa loop to fill with the relevant parameters for heart rate and respiration 
 
Study = NULL
Group = NULL
Trial= NULL
Condition=NULL
  
HR_RR1 = NULL
HR_DET1 = NULL
HR_maxL1 = NULL
HR_L1 = NULL
HR_ENTR1 = NULL
HR_LAM1 = NULL
HR_TT1 = NULL

Resp_RR1 = NULL
Resp_DET1 = NULL
Resp_maxL1 = NULL
Resp_L1 = NULL
Resp_ENTR1 = NULL
Resp_LAM1 = NULL
Resp_TT1 = NULL
n = 1

#Loop to run crqa on all time-series and saving the output in the premade lists 
for (i in files2){
  
  file = read.csv(i)
  
  print(i)
  
  #Extracting study, pair, trial and condition from filenames 
  Study[n] = str_extract(i,"Study+\\d+")
  Group[n] = str_extract(i,"G+\\d+")
  Trial[n] = str_extract(i,"T+\\d+")
  Condition[n] = substr(i, start = 21, stop = 24)
  
#Heart rate
  results = try(crqa(file$HR1S, file$HR2S, delay = hr_delay, embed = hr_embed, radius= hr_radius, normalize=0,rescale=0, mindiagline = 2, minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(results) > 1) {
    HR_RR1[n] <- results$RR
    HR_DET1[n] <- results$DET
    HR_maxL1[n] <- results$maxL
    HR_L1[n] <- results$L
    HR_ENTR1[n] <- results$ENTR
    HR_LAM1[n] <- results$LAM
    HR_TT1[n] <- results$TT
  }
  
#Respiration
    results = try(crqa(file$Resp1S, file$Resp2S, delay = resp_delay, embed = resp_embed, radius= resp_radius, normalize=0,rescale=0, mindiagline = 2, minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(results) > 1) {
    Resp_RR1[n] <- results$RR
    Resp_DET1[n] <- results$DET
    Resp_maxL1[n] <- results$maxL
    Resp_L1[n] <- results$L
    Resp_ENTR1[n] <- results$ENTR
    Resp_LAM1[n] <- results$LAM
    Resp_TT1[n] <- results$TT
  }
  n= n+1
}

#Making a dataframe with the relevant parameters from the crqa 
df2 = data.frame(Study, Group, Trial, Condition, HR_RR = HR_RR1, HR_DET = HR_DET1, HR_maxL = HR_maxL1, HR_L = HR_L1,HR_ENTR = HR_ENTR1, HR_LAM = HR_LAM1, HR_TT = HR_TT1, Resp_RR = Resp_RR1, Resp_DET = Resp_DET1, Resp_maxL = Resp_maxL1, Resp_L = Resp_L1, Resp_ENTR = Resp_ENTR1, Resp_LAM = Resp_LAM1, Resp_TT = Resp_TT1)

#Making two new columns in the dataframe with real data and shuffled data so we can distinguish the data from each other when we put it in the same dataframe  
df1$type = "real" #dataframe from crqa loop
df2$type = "shuffled"

#Binding the real data and the shuffled data into the same dataframe in order to compare them to see whether there is a significant difference
alldata = rbind(df1, df2)

#Making a model in which we see whether the recurrence rate of heart rate differs significantly in the real data from the shuffled data - we do this in order to make sure that we have an actual finding and that the recurrence is not just due to confounding factors (e.g. simply completing the same task)
real_vs_shuffledHR = lmer(HR_RR ~ type + (1|Group), alldata)
summary(real_vs_shuffledHR)
#The model is significant, which means that the recuccerence rate of heart rate in the real data is significantly different from the recurrence rate of heart rate in the shuffled data. The recurrence is therefore not only due to confounding factors 

#Model in which we see whether the recurrence rate of respiration differs significantly in the real data from the shuffled data
real_vs_shuffledResp = lmer(Resp_RR ~ type + (1|Group), alldata)
summary(real_vs_shuffledResp)
#The model is not significantly different, which means that recurrence rate of respiration doesn't differ significantly in the real data compared to the shuffled data. This could mean that the recurrence rate of respiration in the real data could just be due to confounding factors such as completing the same task or just generally breathing. 

```

 
### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair). Tip: Celine will share a commented script
 - Run crqa on all the surrogate pairs and save the output. NB. which delay, embed, radius parameters should you use?
 - Test whether crqa shows a difference between real and surrogate pairs

```{r}
setwd("~/Desktop/Experimental methods 3/Assignment 4/pp")

#Some of the names in the column "condition" were not the same as the others so we just changed their names to be the same 
df1$Condition[which(df1$Condition == "_Tur")] = "Turn"
df1$Condition[which(df1$Condition == "_Syn")] = "Sync"
df1$Condition[which(df1$Condition == "_Con")] = "Conv"

prepro_filelist = list.files(pattern = "pp_Study2") # load the preprocessed files

num=str_extract_all(prepro_filelist, "\\d+") #extract all numbers in the file name (i.e Study?_G?_T?)

df= as.data.frame(t(as.data.frame(num))) #create a transposed (i.e. "tipped") dataframe of these number
rownames(df)<-NULL #Remove rownames.

pairs=as.numeric(as.character(unique(df$V2))) #create pairs, i.e. unique group numbers

SurrogateList <- expand.grid(a = pairs, b = pairs) # expand.grid combines all possible combinations of a given vector
SurrogateList = subset(SurrogateList, a != b) # Ensure that pair 1 (i.e group1) doesn't get paired with it self.


# Per each condition
for (co in c("Sync","Turn","Conv")){ #loop through a list of condintions 
  for (i in 1:nrow(SurrogateList)){ #and loop through all the pairings of the surrogate list
    
  #start recreate filenames from the working directory based on the matching from the surrogatelist
  f1=paste0('Study2_G',SurrogateList$a[i]) #Insert group numeber from Surrogatelist 
  f2=paste0('Study2_G',SurrogateList$b[i])
  
  file1=prepro_filelist[grepl(f1,prepro_filelist) & grepl(co,prepro_filelist)] #insert conditions as well.
  file2=prepro_filelist[grepl(f2,prepro_filelist) & grepl(co,prepro_filelist)]
  
  d1=read.csv(file1) #Load the recreated filenames
  
  d2=read.csv(file2) 
  
  #Ensure both dataframes have the same length
  if (nrow(d1)>nrow(d2)){
    d1=d1[1:nrow(d2),]
  } else {
    d2=d2[1:nrow(d1),]
  }
  
  #creation of surrogate files. (Riccardo's idea of having fun and completely overdoing it ;-) ) 
  #per each 2 pairs there are 4 possible surrogate pairs: 
  #pair1person1 w pair2person2; 
  d3=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp2,HR1=d1$HR1,HR2=d2$HR2,Resp1S=d1$Resp1,Resp2S=d2$Resp2,HR1S=d1$HR1,HR2S=d2$HR2)
  #pair1person2 w pair2person2;
  d4=data.frame(Resp1=d2$Resp1,Resp2=d1$Resp2,HR1=d2$HR1,HR2=d1$HR2,Resp1S=d2$Resp1,Resp2S=d1$Resp2,HR1S=d2$HR1,HR2S=d1$HR2)
  
  #pair1person1 w pair2person1;
  d5=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp1,HR1=d1$HR1,HR2=d2$HR1,Resp1S=d1$Resp1,Resp2S=d2$Resp1,HR1S=d1$HR1,HR2S=d2$HR1)
  #pair1person2 w pair2person2
  d6=data.frame(Resp1=d2$Resp2,Resp2=d1$Resp2,HR1=d2$HR2,HR2=d1$HR2,Resp1S=d2$Resp2,Resp2S=d1$Resp2,HR1S=d2$HR2,HR2S=d1$HR2)
  
  #Make filenames that tells which pairs have been mixed
  file3=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"a_",co,".csv")
  file4=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"b_",co,".csv")
  file5=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"c_",co,".csv")
  file6=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"d_",co,".csv")
  
  #write csv files
  write_csv(d3,file.path("surrogate",file3))
  write_csv(d4,file.path("surrogate",file4))
  write_csv(d5,file.path("surrogate",file5))
  write_csv(d6,file.path("surrogate",file6))
  }
}

```


CRQA loop to see if 

```{r}

setwd("~/Desktop/Experimental methods 3/Assignment 4/pp/surrogate")
files3 = list.files(path = ("~/Desktop/Experimental methods 3/Assignment 4/pp/surrogate"), pattern = "Surrogate")
files3

#Making lists for the crqa loop to fill with the relevant parameters for heart rate and respiration 
HR_RR2 = NULL
HR_DET2 = NULL
HR_maxL2 = NULL
HR_L2 = NULL
HR_ENTR2 = NULL
HR_LAM2 = NULL
HR_TT2 = NULL

Resp_RR2 = NULL
Resp_DET2 = NULL
Resp_maxL2 = NULL
Resp_L2 = NULL
Resp_ENTR2 = NULL
Resp_LAM2 = NULL
Resp_TT2 = NULL
n = 1

#Loop to run crqa on all surrogate time-series and saving the output in the premade lists made above
for (i in files3){
  
  file = read.csv(i)
  
  print(i)
  
#Heart rate
  results = try(crqa(file$HR1S, file$HR2S, delay = hr_delay, embed = hr_embed, radius= hr_radius, normalize=0,rescale=0, mindiagline = 2, minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(results) > 1) {
    HR_RR2[n] <- results$RR
    HR_DET2[n] <- results$DET
    HR_maxL2[n] <- results$maxL
    HR_L2[n] <- results$L
    HR_ENTR2[n] <- results$ENTR
    HR_LAM2[n] <- results$LAM
    HR_TT2[n] <- results$TT
  }
  
#Respiration
    results = try(crqa(file$Resp1S, file$Resp2S, delay = resp_delay, embed = resp_embed, radius= resp_radius, normalize=0,rescale=0, mindiagline = 2, minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(results) > 1) {
    Resp_RR2[n] <- results$RR
    Resp_DET2[n] <- results$DET
    Resp_maxL2[n] <- results$maxL
    Resp_L2[n] <- results$L
    Resp_ENTR2[n] <- results$ENTR
    Resp_LAM2[n] <- results$LAM
    Resp_TT2[n] <- results$TT
  }
  n= n+1
}

#Dataframe with all the crqa parameters for heart recurrence and respiration that we got from the crqa above
df3 = data.frame(HR_RR = HR_RR2, HR_DET = HR_DET2, HR_maxL = HR_maxL2, HR_L = HR_L2,HR_ENTR = HR_ENTR2, HR_LAM = HR_LAM2, HR_TT = HR_TT2, Resp_RR = Resp_RR2, Resp_DET = Resp_DET2, Resp_maxL = Resp_maxL2, Resp_L = Resp_L2, Resp_ENTR = Resp_ENTR2, Resp_LAM = Resp_LAM2, Resp_TT = Resp_TT2)

#Making a new "type" column in the dataframe to distinguish surrogate data from real data 
df3$type = "surrogate"

#In order to merge the dataframe with the real data and the dataframe with the surrogate data, we have to delete study, pair, trial and condition columns from the dataframe with real data as they are not in the dataframe with the surrogate data.
realdf = df1

#Removing study, pair, trial and condition columns from the copied dataframe with the real data 
realdf$Study = NULL
realdf$Group = NULL
realdf$Trial = NULL
realdf$Condition = NULL

#Binding the copied dataframe with real data with the dataframe with surrogate data into one dataframe
alldata1 = rbind(realdf, df3)


#Then we make two models: the first with recurrence rate for heart rate as dependent variable and type as independent variable in order to see whether there is a significant difference between using real data and surrogate data on heart rate recurrence. The second model is only with respiration recurrence rate instead of heart rate recurrence rate. 
real_vs_surrogateHR = lm(HR_RR ~ type, alldata1)
summary(real_vs_surrogateHR)

real_vs_surrogateResp = lm(HR_RR ~ type, alldata1)
summary(real_vs_surrogateResp)

#Both of the models are significant, which means that in both cases the recurrence rate is significantly different when using real data compared to using surrogate data. This indicates that the heart rate and respiration of people from the same pair are significantly more coordinated than the heart rate and respiration of people from different groups (surrogate pairs). 


```

### Testing effects of conditions
 - make a (probably underpowered) mixed model testing effects of the different conditions on heart rate and respiration coordination
 - N.B: would it make sense to include surrogate pairs? and if so how? what would that tell you?

```{r}
#Making trial variable numeric in order to use it as a random effect in the models 
df1$Trial = as.numeric(df1$Trial)

#Model to predict heart rate recurrence rate with condition as predictor to see whether the condition has an effect on heart rate coordination. We use study as a random effect, since we want to correct for possible differences between study 1 and study 2. We also use pair as a random effect because we want to take into consideration that some pairs may generally be more coordinated than others. Lastly, we use trial as a random intercept because people's heart beats may coordinate more in the third trial than the first trial due to aligning with each other.
diff_conHR = lmer(HR_RR ~ condition + (1|Study) + (1 + Trial|Group), df1)
summary(diff_conHR)

#Same model but with respiration instead of heart rate. 
diff_conResp = lmer(Resp_RR ~ condition + (1|Study) + (1 + Trial|Group), df1)
summary(diff_conResp)
```

### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them
 
```{r}

```